---
title: "Homework 2"
author: "mm5951"
date: "`r Sys.Date()`"
output: github_document
---

```{r, include=FALSE}
library(tidyverse)
library(readxl)
```

## Problem 2
First, I import the Mr. Trash Wheel dataset using `read_excel` function (note the library `readxl` has been previously installed, not shown).

Piped into the same code chunk, next, I proceed to read and clean the Mr. Trash Wheel sheet:

* specify the "Mr. Trash Wheel" sheet
* omit non-data entries (by selecting range "A2:N534")
* omit rows that do not include dumpster-specific data
* round the number of sports balls to the nearest integer using `mutate`
* **PENDING - IT DOESN'T WORK, REMAINS AS CHARACTER** convert the result to an integer variable (using as.integer)

```{r}
Mr_trash_df = read_excel("./data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx",
    sheet = "Mr. Trash Wheel",
    range = "A2:N534",) %>% 
  janitor::clean_names() %>% 
  drop_na("dumpster") %>% 
  mutate(sports_balls = round(sports_balls, 0)) %>% 
  transform(sports_balls = as.integer(sports_balls))

typeof("sports_balls")
```

I use a similar process to import, clean, and organize the data for Professor Trash Wheel (same steps **pending - also review NXX number**).

```{r}
Prof_trash_df = read_excel("./data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx",
    sheet = "Professor Trash Wheel",
    range = "A2:N123",) %>%  
  janitor::clean_names() %>% 
  drop_na("dumpster") %>% 
  mutate(sports_balls = round(sports_balls, 0)) %>% 
  transform(sports_balls = as.integer(sports_balls))
```

Next, I and stack the two datasets (that is, combine one on top of the other) using the `bind_rows` function. To do so, first I generate an "origin" column, containing the prefix "M" for Mr Trash and "P" for the Professor dataframes. In doing so, I use the function `mutate` to generate a "M" or "P" containing columns, and then merge it using `unite` with the corresponding "dumpster" number, creating an "Id" column.

```{r}
Mr_trash2_df = Mr_trash_df %>% 
  mutate(origin = "M", .before = "dumpster")  %>% 
  unite('Id', origin:dumpster, sep = "", remove = FALSE)

Prof_trash2_df = Prof_trash_df %>% 
  mutate(origin = "P", .before = "dumpster")  %>% 
  unite('Id', origin:dumpster, sep = "", remove = FALSE)

stacked_trash_df = bind_rows(Mr_trash2_df, Prof_trash2_df) %>% 
  janitor::clean_names()
```

The resulting merged dataset contains `R col()`XX observations, n=453 from Mr Trash and n=71 from Professor trash.

In "Mr. Trash Wheel dataset", the key variables are listed as: `r names(Mr_trash_df)`. Mr. Trash Wheel data contains `r nrow(Mr_trash_df)` observations of dumpsters, and collects data about `r ncol(Mr_trash_df)` related variables. Notably, some key results:

* Mr. Trash collected a total of `r sum(Mr_trash_df$weight_tons)` tons of dumpster over the entire study period.
* In 2020, he collected a total of `r sum(select(filter(Mr_trash_df$sports_balls, Mr_trash_df$year == 2020), total))`sports balls.

** PENDING SOME FILTERING ERROR FOR THE RESULTS***


## Problem 2
** PENDING**
