---
title: "Homework 2"
author: "mm5951"
date: "`r Sys.Date()`"
output: github_document
---

```{r, include=FALSE}
library(tidyverse)
library(readxl)
```

## Problem 1
Solutions provided in the p8105 course page.

## Problem 2
First, I import the Mr. Trash Wheel dataset using `read_excel` function (note the library `readxl` has been previously installed, not shown).

Piped into the same code chunk, I proceed to read and clean the Mr. Trash Wheel sheet:

* specify the "Mr. Trash Wheel" sheet
* omit non-data entries (by selecting range "A2:N549")
* omit rows that do not include dumpster-specific data
* round the number of sports balls to the nearest integer using `mutate`
* convert the result to an integer variable using the `transform` and `as.integer`function

```{r}
Mr_trash_df = read_excel("./data/Trash-Wheel-Collection-Data.xlsx",
    sheet = "Mr. Trash Wheel",
    range = "A2:N549",) %>% 
  janitor::clean_names() %>% 
  drop_na("dumpster") %>% 
  mutate(sports_balls = round(sports_balls, 0)) %>% 
  transform(sports_balls = as.integer(sports_balls)) 
```

Finally, I use the `str()` function to compactly display the structure of the Mr_trash_df dataframe and check the integer nature of the "sports_balls" variable.

```{r}
  str(Mr_trash_df$"sports_balls")
```

I use a similar process to import, clean, and organize the data for Professor Trash Wheel (same steps and functions as described above). Note the variable "year" is converted into a character using `as.character` and "sports_balls" is newly created and transformed into a integer variable in order to allow for merging in a later step.

```{r, warning = FALSE}
Prof_trash_df = read_excel("./data/Trash-Wheel-Collection-Data.xlsx",
    sheet = "Professor Trash Wheel",
    range = "A2:M96",) %>%  
  janitor::clean_names() %>% 
  drop_na("dumpster") %>% 
  mutate(sports_balls = "NA", .after = "chip_bags") %>% 
  transform(year = as.character(year)) %>% 
  transform(sports_balls = as.integer(sports_balls)) 
```

Next, I and stack the two datasets (that is, combine one on top of the other) using the `bind_rows` function. To do so, first I generate an "origin" column, containing the prefix "M" for Mr Trash and "P" for the Professor dataframes. In doing so, I use the function `mutate` to generate a "M" or "P" containing columns, and then merge it using `unite` with the corresponding "dumpster" number, creating an "Id" column.

```{r}
Mr_trash2_df = Mr_trash_df %>% 
  mutate(origin = "M", .before = "dumpster")  %>% 
  unite('Id', origin:dumpster, sep = "", remove = FALSE)

Prof_trash2_df = Prof_trash_df %>% 
  mutate(origin = "P", .before = "dumpster")  %>% 
  unite('Id', origin:dumpster, sep = "", remove = FALSE)

stacked_trash_df = bind_rows(Mr_trash2_df, Prof_trash2_df) %>% 
  janitor::clean_names()
```

The resulting merged dataset contains `r nrow(stacked_trash_df)` observations, n=547 from Mr Trash and n=94 from Professor trash.

In "Mr. Trash Wheel dataset", the key variables are listed as: `r names(Mr_trash_df)`. Mr. Trash Wheel data contains `r nrow(Mr_trash_df)` observations of dumpsters, and collects data about `r ncol(Mr_trash_df)` related variables. Notably, some key results are:

* Over the entire study period, Prof. Trash collected a total of `r sum(Prof_trash_df$weight_tons)` tons of dumpster.
* In 2020 alone, he collected a total of `r sum(pull(filter(Mr_trash_df, year == 2020), sports_balls))` sports balls.


## Problem 3

First, I import the **"pols-month.csv" dataset** (part of "FiveThirtyEight data") using the `read_csv` function. More specifically, the following functions are used:

* To clean names, the `janitor::clean_names()` function;
* To break the date "mon" variable into individual year, month, and day, the `separate()` function;
* To coerce into integer data types, the `as.integer()` function;
* To replace month number with month name, as well as create a president variable taking values "gop" and "dem", the `mutate()` function;
* To remove the "prez_dem", "prez_gop" and "day" variables, the `select()`function.
* To arrange according to year and month, the `arrange()` function.

```{r}
polsmonth_df = read_csv("./Data/pols-month.csv", show_col_types = FALSE) %>%
  janitor::clean_names() %>%
  separate(mon, into = c("year", "month", "day")) %>%
  arrange(year, as.integer(month)) %>%
  mutate(year = as.integer(year),
         month = month.name[as.integer(month)],
         president = if_else(prez_dem == 1, "dem", "gop")
         ) %>%
  select(-day, -prez_dem, -prez_gop)
```

Next, I proceed similarly as described above for the **"snp.csv"** and **"unemployment.csv" datasets**.

```{r}
snp_df = read_csv("./Data/snp.csv", show_col_types = FALSE) %>%
  janitor::clean_names() %>%
  separate(date, into = c("month", "day", "year")) %>%
  mutate(year = as.integer(year),
         month = as.integer(month),
         year = ifelse(year >= 23, year + 1900, year + 2000)
         ) %>%
  relocate(year) %>%
  arrange(year, month) %>%
  mutate(month = month.name[as.integer(month)]) %>%
  select(-day)
```

For the "unemployment.csv" dataset, some additional steps are taken:

* Switch from “wide” to “long” format using `pivot_longer()`function; 
* Renaming some of the key variables so they have the same name and take the same values, using `mutate()` and `as.integer()`.

```{r}
unemployment_df = read_csv("./Data/unemployment.csv", show_col_types = FALSE) %>%
  pivot_longer("Jan":"Dec",
        names_to = "month",
        values_to = "unemployment"
        ) %>%
  janitor::clean_names() %>%
  mutate( month = month.name[match(month, month.abb)],
          year = as.integer(year)
         )
```

Then I proceed to **join the datasets**, first by merging "snp_df" into "polsmonth_df", and then merging "unemployment_df" into the result. To do so, I use the `full_join()` function.

```{r, message = FALSE}
joint_df = full_join(polsmonth_df, snp_df)
joint2_df = full_join(joint_df, unemployment_df)
```

Finally, I describe the contents of the overall "joint2_df" dataset, as well as the dimension, range of years, and names of key variables of each separate original ones.

* The **"pols-month.csv" dataset** entails:
    -   The dataset has `r nrow(polsmonth_df)` observations and `r ncol(polsmonth_df)` variables (columns).
    -   The years range from `r min(polsmonth_df$year)` to `r max(polsmonth_df$year)`.
    -   The key variables are listed as: `r names(polsmonth_df)`.

* The **"snp.csv"** dataset entails:
    -   The dataset has `r nrow(snp_df)` observations and `r ncol(snp_df)` variables (columns).
    -   The years range from `r min(snp_df$year)` to `r max(snp_df$year)`.
    -   The key variables are listed as:  `r names(snp_df)`.

* The **"unemployment.csv"** dataset entails:
    -   The dataset has `r nrow(unemployment_df)` observations and `r ncol(unemployment_df)` variables (columns).
    -   The years range from `r min(unemployment_df$year)` to `r max(unemployment_df$year)`.
    -   The key variables are listed as:  `r names(unemployment_df)`.

* The final **"joint2_df"** dataset, which is the result of a merging from the three described above, entails:
    -   The dataset has `r nrow(joint2_df)` observations and `r ncol(joint2_df)` variables (columns).
    -   The years range from `r min(joint2_df$year)` to `r max(joint2_df$year)`
    -   The key variables are listed as:  `r names(joint2_df)`.

